---
title: "Feasibility Summary"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
---

```{r setup, include=FALSE}
# Default taken from R notebook behavior: when knitting, wd will always be location of notebook
base_dir <- ".."
Sys.setenv(PEDSNET_SKIP_REQUEST_EXECUTION = 1L)
tryCatch(source("../site/run.R"),
  error = function(e) message(e)
) # May not be able to make db connection

# Set to "local" to read data from ../results, or another value to read from db
data_source <-
  if_else(config("execution_mode") == "distribution",
    "local",
    "not_local"
  )

require(tidyr)
require(knitr)
require(kableExtra)
require(stringr)
require(tibble)
require(ggplot2)
require(DT)
require(gt)
require(table1)

get_results <- function(tbl_name) {
  if (data_source == "local") {
    rslt <- read_csv(paste0("../results/", tbl_name, ".csv"))
  } else {
    rslt <- results_tbl(tbl_name) %>% collect()
  }
  rslt
}

prettify_kable <- function(data, title = "") {
  rslt <-
    data %>% kable(
      digits = 4L,
      caption = title,
      format.args = list(big.mark = ",")
    )
  if (knitr::is_html_output()) {
    rslt <- rslt %>%
      kable_styling(bootstrap_options = c("striped", "hover")) %>%
      column_spec(1L, bold = TRUE, border_right = TRUE)
  }
  rslt
}

cohort <- results_tbl("cohort_post_attrition")
# Manually mask the site name
cohort <- cohort %>%
  left_join(
    results_tbl("xwalk_site"),
    by = "site") %>%
  select(-site) %>%
  mutate(site = as.character(seq_id)) %>%
  select(-seq_id)
cohort <- filter_no_treatment(cohort, treatment_window = 30L)
measurements_fixed <- make_manual_fixes(results_tbl("measurement"))
# Manually mask the site name
measurements_fixed <- measurements_fixed %>%
  left_join(
    results_tbl("xwalk_site"),
    by = "site") %>%
  select(-site) %>%
  mutate(site = as.character(seq_id)) %>%
  select(-seq_id)

redaction_str <- "[REDACTED]"
low_count_threshold <- 10L
round_to <- 5L
```

## Notices

### Feasibility Query

The data presented here are the results of a feasbility evaluation performed by
the PEDSnet data and analytic coordinating center.  Because feasibility queries
require efficient completion, they entail less consultation regarding study
design, resource allocation, and data quality analysis than full study
execution, feasibility results may incompletely predict observations made
during the study.

Counts less than `r low_count_threshold` have been redacted. If in a group the
total redacted is less than `r low_count_threshold`, the next smallest number
was also redacted to protect against subtracting from the total. Counts have
also been rounded to the nearest `r round_to` to protect against small
differences in similar analyses or re-runs.

### Confidentiality

Please treat the contents of this report as confidential, and do not
redistrubute beyond PEDSnet staff and the requester.

## Request Information

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
tribble(
  ~Item, ~Value,
  "Title", config("title"),
  "Requester", config("requester_name"),
  "Sponsoring Member", config("requester_site"),
  "CDM Version", as.character(config("cdm_version")),
  "Request Version", as.character(config("req_version"))
) %>%
  prettify_kable()
```

## Results

The cohort comprises `r summarise(cohort, ct = n_distinct(person_id)) %>% pull(ct)`
patients from `r summarise(cohort, ct = n_distinct(site)) %>% pull(ct)` sites.

### Attrition

```{r attrition, echo=FALSE, message=FALSE, warning=FALSE}
# DBI doesn't have a good way to interrogate tables in a specified schema,
# so you have to list the step logs you want to summarise (e.g. DCC + remote
# sites).
step_log <- list(config("qry_site")) %>%
  map(.f = function(x) {
    get_results(paste0(x, "_steps"))
  }) %>%
  reduce(.f = dplyr::union)

step_log %>%
  arrange(stamp) %>%
  select(cohort, persons) %>%
  redact_and_round(
    columns = "persons",
    threshold = low_count_threshold,
    round_to = round_to
  ) %>%
  mutate(percent_prior = round(100L * persons /
    coalesce(lag(persons), persons), 2L)) %>%
  replace_na_subset(
    columns = c("persons", "percent_prior"),
    redaction_val = redaction_str
  ) %>%
  prettify_kable()
```

### Site stratification

This stratification comprises the final included cohort.

```{r site, echo=FALSE, message=FALSE, warning=FALSE}
cohort %>%
  group_by(site) %>%
  summarize(persons = n_distinct(person_id)) %>%
  collect() %>%
  redact_and_round(
    columns = "persons",
    threshold = low_count_threshold,
    round_to = round_to
  ) %>%
  mutate(
    percent_total =
      round(100L * persons / sum(persons, na.rm = TRUE), 2L)
  ) %>%
  replace_na_subset(
    columns = c("persons", "percent_total"),
    redaction_val = redaction_str
  ) %>%
  prettify_kable()
```

### Diagnoses over Time
> NOTE! This does not adjust for a potentially changing number of registered
patients in the cohort over time, or changing number of registered patient per
PEDSnet site, it is a simple crude count

```{r time, echo=FALSE, message=FALSE, warning=FALSE}
over_time <- cohort %>%
  mutate(
    start_year = year(dx_hyperthyroid_date),
    age_at_dx = (dx_hyperthyroid_date - birth_datetime)
  ) %>%
  select(person_id, start_year, age_at_dx, site) %>%
  collect() %>%
  # Ensure x-axis has full years not e.g. 2015.5
  mutate(start_year = as.Date(as.character(start_year), format = "%Y"))

overall <- over_time %>%
  count(start_year) %>%
  mutate(site = "Overall")

by_site <- over_time %>%
  count(start_year, site)

bind_rows(overall, by_site) %>%
  redact_and_round(
    columns = "n",
    groupers = "site",
    threshold = low_count_threshold,
    round_to = round_to
  ) %>%
  ggplot(aes(x = start_year, y = n, color = site)) +
  geom_line() +
  ggtitle("Raw count of new Graves Diagnoses") +
  xlab("Year") +
  ylab("Count of persons")
```

### Age at diagnosis
```{r age, echo=FALSE, message=FALSE, warning=FALSE}
over_time %>%
  mutate(age_in_years = floor(as.integer(str_split_i(
    age_at_dx,
    pattern = " ", i = 1L
  )) / 365.25)) %>%
  count(age_in_years) %>%
  redact_and_round(
    columns = "n",
    groupers = "age_in_years",
    threshold = low_count_threshold,
    round_to = round_to
  ) %>%
  ggplot(aes(x = age_in_years, y = n)) +
  geom_bar(stat = "identity") +
  ggtitle("Age at new Graves Diagnosis") +
  xlab("Age in Years") +
  ylab("Count of persons")
```

### Demographics

This stratification comprises the final included cohort. Demographic
categories from the typical PHS398 set not listed here reflect zero
cohort members in that category.

```{r demog, echo=FALSE, message=FALSE, warning=FALSE}
c_names <- vocabulary_tbl("concept") %>%
  filter(domain_id %in% c("Gender", "Race", "Ethnicity")) %>%
  select(concept_id, concept_name) %>%
  collect()

demog <- cdm_tbl("person") %>%
  semi_join(distinct(cohort, person_id),
    by = "person_id", copy = TRUE
  ) %>%
  collect() %>%
  inner_join(select(c_names, concept_id, sex = concept_name),
    by = c("gender_concept_id" = "concept_id")
  ) %>%
  inner_join(select(c_names, concept_id, race = concept_name),
    by = c("race_concept_id" = "concept_id")
  ) %>%
  inner_join(select(c_names, concept_id, hispanic = concept_name),
    by = c("ethnicity_concept_id" = "concept_id")
  ) %>%
  select(person_id, sex, race, hispanic)

h_cts <- demog %>%
  group_by(sex, hispanic) %>%
  summarise(persons = n()) %>%
  redact_and_round(
    columns = "persons",
    groupers = c("sex", "hispanic"),
    threshold = low_count_threshold,
    round_to = round_to,
    redaction_val = redaction_str
  ) %>%
  pivot_wider(names_from = sex, values_from = persons) %>%
  rename(group = hispanic)
r_cts <- demog %>%
  group_by(sex, race) %>%
  summarise(persons = n()) %>%
  redact_and_round(
    columns = "persons",
    groupers = c("sex", "race"),
    threshold = low_count_threshold,
    round_to = round_to
  ) %>%
  pivot_wider(names_from = sex, values_from = persons) %>%
  rename(group = race)

prettify_kable(r_cts)
prettify_kable(h_cts)
```

### Payer

Primary payer plan class. If there is more than one class, it is categorized as
"multiple". We use the payer information recorded closest to the presenting
visit, as we had high missingness looking just at the presenting visit.
```{r payer, echo=FALSE, message=FALSE, warning=FALSE}
# Calculate the time between the visit with the payer info and the presenting
# visit
all_visits <- (cohort %>%
  select(person_id, dx_hyperthyroid_date, site)) %>%
  inner_join((
    results_tbl("visit_occurrence") %>%
      select(person_id, visit_occurrence_id, visit_start_date)
  ), by = "person_id") %>%
  mutate(time_to_payer = visit_start_date - dx_hyperthyroid_date)

payer_count <- all_visits %>%
  inner_join(results_tbl("visit_payer"), by = "visit_occurrence_id") %>%
  group_by(person_id) %>%
  slice_min(order_by = tibble(abs(time_to_payer)), with_ties = TRUE) %>%
  ungroup() %>%
  group_by(person_id) %>%
  summarize(
    plan_count = n_distinct(plan_class),
    plan = min(plan_class),
    time_to_payer = min(time_to_payer)
  )

payer_class <- cohort %>%
  left_join(payer_count, by = "person_id") %>%
  mutate(category = if_else(plan_count > 1L, "Multiple", plan)) %>%
  count(category) %>%
  arrange(desc(n)) %>%
  collect()

assertthat::assert_that(
  sum(payer_class[["n"]]) == (cohort %>% distinct_ct()),
  msg = "Does not contain one row per person in cohort"
)

prettify_kable(
  payer_class %>%
    redact_and_round(
      columns = "n",
      threshold = low_count_threshold,
      round_to = round_to,
      redaction_val = redaction_str
    )
)

time_to_payer <- payer_count %>%
  summarize(
    mean_time_to_payer = mean(time_to_payer),
    sd_time_to_payer = sd(time_to_payer)
  ) %>%
  mutate_if(is.numeric, round, 1L) %>%
  collect()
```
The mean (sd) of time between the presenting visit and a visit with insurance
information is `r time_to_payer[["mean_time_to_payer"]]` (`r time_to_payer[["sd_time_to_payer"]]`).

## Specification

### Feasibility Notes

#### Measurements at presentation
What proportion of the final cohort had each measurement up to 7 days after
the diagnosing encounter? ex -10 would count anyone with a code in that
codeset between -10 and +7 days of the diagnosing encounter, and -20 would count
anyone with a code in that codeset between -20 and +7 days of the diagnosing
encounter. 
```{r measurements, echo=FALSE, message=FALSE, warning=FALSE}
results_df <- proportion_over_time(
  c(
    "mx_bmi",
    "mx_diastolic",
    "mx_systolic",
    "mx_tsh",
    "mx_heart_rate",
    "mx_tsi",
    "mx_trab",
    "mx_free_t4",
    "mx_total_t3",
    "mx_anti_tpo",
    "mx_anti_tg",
    "mx_dx_vitamin_d",
    "mx_covid"
  ),
  domain = "measurement"
)
plot_all_codesets(results_df)
```

Measurement records can have numeric values (e.g. 0.01) or concepts (e.g.
Negative, Clear) associated with them. It is also possible that there is no
value associated with a measurement, or the value failed to parse (in some cases
<0.01). The following plot demonstrates the types of values available for each
codeset of interest.
```{r result_type, echo=FALSE, message=FALSE, warning=FALSE}
plot_result_type_panel(results_df)
```

#### Missingness
```{r diagnoses, echo=FALSE, message=FALSE, warning=FALSE}
time_between_dx <- cohort %>%
  mutate(diff = dx_graves_date - dx_hyperthyroid_date) %>%
  summarize(mean = mean(diff, na.rm = TRUE), sd = sd(diff, na.rm = TRUE)) %>%
  mutate_if(is.numeric, round, 1L) %>%
  collect()

treatment <- cohort %>%
  filter(is.na(dx_graves_date)) %>%
  count() %>%
  collect() %>%
  redact_and_round(columns = "n")
```
How much missingness is there in each codeset?
We can look at the initial diagnosis, which can include a general diagnosis of
hyperthyroidism, or we can look specifically at the first Graves diagnosis.

The average (sd) time between the two diagnoses is
`r time_between_dx[["mean"]]` (`r time_between_dx[["sd"]]`).

Around `r treatment` people had a thyroidectomy on or before their Graves
diagnosis, or were treated with anti-thyroid medication more than 30 days before
their Graves diagnosis, so for those people we will only use the hyperthyroidism
diagnosis.

```{r missing, echo=FALSE, message=FALSE, warning=FALSE}
missing_df <- describe_missing_by_site(
  c(
    "mx_bmi",
    "mx_diastolic",
    "mx_systolic",
    "mx_heart_rate",
    "mx_tsh",
    "mx_tsi",
    "mx_trab",
    "mx_free_t4",
    "mx_total_t3",
    "mx_anti_tpo",
    "mx_anti_tg",
    "mx_dx_vitamin_d"
  ),
  cohort,
  measurements_fixed,
  min_days = -20L,
  max_days = 7L,
  redact = TRUE
)

missing_df %>%
  prettify_kable(title = "Missingness by Site")
```

#### Standardizing units
What units do different codesets have?

```{r units, fig.width = 12, fig.height = 12, echo=FALSE, message=FALSE, warning=FALSE}
measurements <-
  c(
    "mx_bmi",
    "mx_diastolic",
    "mx_systolic",
    "mx_heart_rate",
    "mx_tsh",
    "mx_tsi",
    "mx_trab",
    "mx_free_t4",
    "mx_total_t3",
    "mx_anti_tpo",
    "mx_anti_tg",
    "mx_dx_vitamin_d"
  )

all_thresholds <- c()
for (measurement in measurements) {
  unit_data <-
    get_numeric_unit_data(results_tbl(measurement), cohort, measurements_fixed)
  plot(plot_unit_histogram(unit_data, redact = TRUE) +
    ggtitle(paste0(measurement, " measurement values")))

  operator_counts <- unit_data %>%
    count(operator_concept_name, name = "operator_count") %>%
    mutate(codeset = measurement, total = sum(operator_count))

  thresholds <- unit_data %>%
    filter(
      !is.na(operator_concept_name),
      !operator_concept_id %in% c(4172703L, 0L)
    ) %>%
    count(operator_concept_name, value_as_number, name = "threshold_count")

  threshold_description <- operator_counts %>%
    left_join(thresholds, by = "operator_concept_name")

  all_thresholds <- append(all_thresholds, list(threshold_description))
}

bind_rows(all_thresholds) %>%
  arrange(codeset, desc(operator_count), desc(threshold_count)) %>%
  mutate(
    percent =
      paste0("(", round(100L * operator_count / total, 0L), "%)")
  ) %>%
  mutate(
    threshold_percent =
      paste0("(", round(100L * threshold_count / total, 0L), "%)")
  ) %>%
  select(
    codeset,
    operator_concept_name,
    percent,
    value_as_number,
    threshold_percent
  ) %>%
  group_by(codeset) %>%
  gt(row_group.sep = " ")
```

Takeaways:

* No adjustments are needed to the units for bmi, diastolic, systolic, heart
rate, anti-tpo, anti-tg
* Some measurements have incorrect mappings of source values, but the values
themselves do not need to be adjusted
* TSI has two units ("percent baseline" and "International unit per liter") - do
we want to report these two units separately? At one site representing about 8%
of TSI measurements, there is no unit source value. Do we want to assume these
are TSI index?
* Some measurements have units that need to be mapped to the standard unit e.g.
"TSI index" * 100 to convert to "percent baseline", and total T3
"Nanogram per milliliter" * 100 to convert to "Nanogram per deciliter"
* Free T4 is an example some labs reporting values >10 e.g. for the
"No Information" unit concept, we see a cluster of values at 10. If some labs
only allow values up to 10, and other labs allow values up to 50, it would
impact the overall average value we see.
* 46% of TSH values have a < operator, but all with values less than 0.03. It
might skew the reported value, but we could set these values equal to their
threshold. We could consider reporting the outcome variable to be
"proportion abnormal" rather than reporting the continuous values

#### Reference Ranges
For anti-tpo and anti-tg, we discussed using the lab provided references
```{r ref_ranges, echo=FALSE, message=FALSE, warning=FALSE}
anti_tpo <-
  find_closest_2_dates(results_tbl("mx_anti_tpo"),
                       cohort,
                       measurements_fixed,
                       "measurement") %>%
  filter(!is.na(measurement_id))

anti_tpo_totals_site <- anti_tpo %>% count(site, name = "total")

anti_tpo %>%
  count(site, range_high) %>%
  left_join(anti_tpo_totals_site, by = "site") %>%
  redact_and_round(
    columns = c("n", "total"),
    groupers = c("site"),
    threshold = low_count_threshold,
    round_to = round_to
  ) %>%
  mutate(pct = round(100 * n / total, 0),
         label = if_else(is.na(pct), "[REDACTED]", paste0(n, " (", pct, "%)"))
         ) %>%
  select(range_high, site, label) %>%
  pivot_wider(names_from = "site",
              values_from = "label",
              values_fill = "0") %>%
  arrange(range_high) %>%
  gt()

anti_tg <-
  find_closest_2_dates(results_tbl("mx_anti_tg"),
                       cohort,
                       measurements_fixed,
                       "measurement") %>%
  filter(!is.na(measurement_id))

anti_tg_totals_site <- anti_tg %>% count(site, name = "total")

anti_tg %>%
  count(site, range_high) %>%
  left_join(anti_tg_totals_site, by = "site") %>%
  redact_and_round(
    columns = c("n", "total"),
    groupers = c("site"),
    threshold = low_count_threshold,
    round_to = round_to
  ) %>%
  mutate(pct = round(100 * n / total, 0),
         label = if_else(is.na(pct), "[REDACTED]", paste0(n, " (", pct, "%)"))
         ) %>%
  select(range_high, site, label) %>%
  pivot_wider(names_from = "site",
              values_from = "label",
              values_fill = "0") %>%
  arrange(range_high) %>%
  gt() %>%
  tab_header(title = "Anti-tg lab provided reference 'range_high'")
```

## End
